---
title: "COVID-19 Trend in BC"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a breakdown of every confirmed or presumed COVID-19 case in BC over time. (source: https://www.ctvnews.ca/health/coronavirus/tracking-every-case-of-covid-19-in-canada-1.4852102)

```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
covidbc.data = read.csv("covid.csv")
#covidbc.data
covidbc.data$BCDate <- as.Date(covidbc.data$BCDate)
#sapply(covidbc.data, class)

totalcases <- sum(covidbc.data$BCCases)
totaldeaths <- sum(covidbc.data$BCDeaths)

covidbc.bcsub <- subset(covidbc.data, select=c("BCDate","BCCases","BCDeaths"))

covidbc.long <- covidbc.bcsub %>%
  gather("Stat", "Value", -BCDate)
#covidbc.long
```

Total cases as of `r format(Sys.time(), '%B %d, %Y')`: `r totalcases`  
Total deaths: `r totaldeaths`

## Plot

```{r, echo=FALSE}
library(ggplot2)

ggplot(data=covidbc.long, aes(x = BCDate, y = Value, fill = Stat)) + geom_col(position = "dodge") + guides(fill=guide_legend(title=NULL)) + xlab("Date") + ylab("New Cases")


```

## Italy

Let's compare this trend with Italy.

The population of Vancouver Metro is ~2.5M

The population of Italy is ~60.5M

Which gives us a ratio of `r 2.5 / 60.5` Vancouver to Italy

Using stats from https://www.worldometers.info/coronavirus/country/italy/, let's see if we can find a time that Italy had a similar Daily New Case level.

```{r, echo=FALSE}
covidbc.data$ItalyCasesBC <- covidbc.data$ItalyCases*0.0413223
#covidbc.data$ItalyCasesBC

#covidbc.idata <- data.frame(casedate,cases,deaths,icasesbc)
covidbc.idata <- subset(covidbc.data, select=c("BCDate","BCCases","ItalyCasesBC","BCDeaths"))

covidbc.ilong <- covidbc.idata %>%
  gather("Stat", "Value", -BCDate)

ggplot(data=covidbc.ilong, aes(x = BCDate, y = Value, fill = Stat)) + geom_col(position = "dodge") + guides(fill=guide_legend(title=NULL)) + xlab("Date") + ylab("New Cases")
```

## Log Trend March 16th, 2020

Let's take the raw data, perform a log transform, and then a linear fit
We can then try to predict future numbers

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#covidbc.data$BCCasesLog <- log10(covidbc.data$BCCases)
covidbc.middle <- subset(covidbc.data, DayFirstCase > 39 & DayFirstCase < 50)

covidbc.middle$BCCasesLog <- log10(covidbc.middle$BCCases)
#covidbc.middle

covidbc.logdata <- subset(covidbc.middle, select=c("DayFirstCase","BCCasesLog"))

covidbc.logdata$BCCasesLog[is.infinite(covidbc.logdata$BCCasesLog)] <- NA

modelbc <- lm(formula = BCCasesLog ~ DayFirstCase, data = covidbc.logdata)

new.dayfirstcase <- data.frame(DayFirstCase = 50:60)
new.dayfirstcase$BCCasesLog = predict(modelbc, newdata = new.dayfirstcase)

p <- ggplot(covidbc.logdata, aes(DayFirstCase, BCCasesLog)) +
  geom_point()
p + geom_smooth(method = "lm")

covidbc.data$BCCasesPredicted = covidbc.data$BCCases


covidbc.data[50:60, "BCCasesPredicted"] = 10^(new.dayfirstcase$BCCasesLog)

#covidbc.logdata[50:60, "BCCasesLog"] = new.dayfirstcase$BCCasesLog

#covidbc.data$BCCasesPredicted <-10^(covidbc.logdata$BCCasesLog)

#covidbc.data$BCCasesPredicted

covidbc.bcsubpred <- subset(covidbc.data, select=c("BCDate","BCCasesPredicted","BCDeaths", "ItalyCasesBC"))

covidbc.longpred <- covidbc.bcsubpred %>%
  gather("Stat", "Value", -BCDate)
```

## 

```{r, echo= FALSE}
ggplot(data=covidbc.longpred, aes(x = BCDate, y = Value, fill = Stat)) + geom_col(position = "dodge") + guides(fill=guide_legend(title=NULL)) + xlab("Date") + ylab("New Cases")

#covidbc.bcsubpred

```

Given this, I would predict `r round(covidbc.bcsubpred$BCCasesPredicted[50])` new cases to be announced tomorrow (March 17th)

## Log Trend March 17th, 2020

Today it was announced there were 83 new cases  
And 3 new deaths  
Here is the new predictions

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#covidbc.data$BCCasesLog <- log10(covidbc.data$BCCases)
covidbc.middle <- subset(covidbc.data, DayFirstCase > 39 & DayFirstCase < 51)

covidbc.middle$BCCasesLog <- log10(covidbc.middle$BCCases)
#covidbc.middle

covidbc.logdata <- subset(covidbc.middle, select=c("DayFirstCase","BCCasesLog"))

covidbc.logdata$BCCasesLog[is.infinite(covidbc.logdata$BCCasesLog)] <- NA

modelbc <- lm(formula = BCCasesLog ~ DayFirstCase, data = covidbc.logdata)

new.dayfirstcase <- data.frame(DayFirstCase = 51:60)
new.dayfirstcase$BCCasesLog = predict(modelbc, newdata = new.dayfirstcase)

p <- ggplot(covidbc.logdata, aes(DayFirstCase, BCCasesLog)) +
  geom_point()
p + geom_smooth(method = "lm")

covidbc.data$BCCasesPredicted = covidbc.data$BCCases


covidbc.data[51:60, "BCCasesPredicted"] = 10^(new.dayfirstcase$BCCasesLog)

covidbc.bcsubpred <- subset(covidbc.data, select=c("BCDate","BCCasesPredicted","BCDeaths", "ItalyCasesBC"))

covidbc.longpred <- covidbc.bcsubpred %>%
  gather("Stat", "Value", -BCDate)
```

## 

```{r, echo= FALSE}

ggplot(data=covidbc.longpred, aes(x = BCDate, y = Value, fill = Stat)) + geom_col(position = "dodge") + guides(fill=guide_legend(title=NULL)) + xlab("Date") + ylab("New Cases")

#covidbc.bcsubpred

```

Given this, I would predict `r round(covidbc.bcsubpred$BCCasesPredicted[51])` new cases to be announced tomorrow (March 18th)

## Log Trend March 18th, 2020

Today it was announced there were 45 new cases  
And 0 new deaths  
Here is the new predictions

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#covidbc.data$BCCasesLog <- log10(covidbc.data$BCCases)
covidbc.middle <- subset(covidbc.data, DayFirstCase > 39 & DayFirstCase < 52)

covidbc.middle$BCCasesLog <- log10(covidbc.middle$BCCases)
#covidbc.middle

covidbc.logdata <- subset(covidbc.middle, select=c("DayFirstCase","BCCasesLog"))

covidbc.logdata$BCCasesLog[is.infinite(covidbc.logdata$BCCasesLog)] <- NA

modelbc <- lm(formula = BCCasesLog ~ DayFirstCase, data = covidbc.logdata)

new.dayfirstcase <- data.frame(DayFirstCase = 52:60)
new.dayfirstcase$BCCasesLog = predict(modelbc, newdata = new.dayfirstcase)

p <- ggplot(covidbc.logdata, aes(DayFirstCase, BCCasesLog)) +
  geom_point()
p + geom_smooth(method = "lm")

covidbc.data$BCCasesPredicted = covidbc.data$BCCases


covidbc.data[52:60, "BCCasesPredicted"] = 10^(new.dayfirstcase$BCCasesLog)

covidbc.bcsubpred <- subset(covidbc.data, select=c("BCDate","BCCasesPredicted","BCDeaths", "ItalyCasesBC"))

covidbc.longpred <- covidbc.bcsubpred %>%
  gather("Stat", "Value", -BCDate)
```

## 

```{r, echo= FALSE}

ggplot(data=covidbc.longpred, aes(x = BCDate, y = Value, fill = Stat)) + geom_col(position = "dodge") + guides(fill=guide_legend(title=NULL)) + xlab("Date") + ylab("New Cases")

#covidbc.bcsubpred

```

Given this, I would predict `r round(covidbc.bcsubpred$BCCasesPredicted[52])` new cases to be announced tomorrow (March 19th)

## Log Trend March 19th, 2020

Today it was announced there were 40 new cases  
And 1 new deaths  
Here is the new predictions

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#covidbc.data$BCCasesLog <- log10(covidbc.data$BCCases)
covidbc.middle <- subset(covidbc.data, DayFirstCase > 39 & DayFirstCase < 53)

covidbc.middle$BCCasesLog <- log10(covidbc.middle$BCCases)
#covidbc.middle

covidbc.logdata <- subset(covidbc.middle, select=c("DayFirstCase","BCCasesLog"))

covidbc.logdata$BCCasesLog[is.infinite(covidbc.logdata$BCCasesLog)] <- NA

modelbc <- lm(formula = BCCasesLog ~ DayFirstCase, data = covidbc.logdata)

new.dayfirstcase <- data.frame(DayFirstCase = 53:60)
new.dayfirstcase$BCCasesLog = predict(modelbc, newdata = new.dayfirstcase)

p <- ggplot(covidbc.logdata, aes(DayFirstCase, BCCasesLog)) +
  geom_point()
p + geom_smooth(method = "lm")

covidbc.data$BCCasesPredicted = covidbc.data$BCCases


covidbc.data[53:60, "BCCasesPredicted"] = 10^(new.dayfirstcase$BCCasesLog)

covidbc.bcsubpred <- subset(covidbc.data, select=c("BCDate","BCCasesPredicted","BCDeaths", "ItalyCasesBC"))

covidbc.longpred <- covidbc.bcsubpred %>%
  gather("Stat", "Value", -BCDate)
```

## 

```{r, echo= FALSE}

ggplot(data=covidbc.longpred, aes(x = BCDate, y = Value, fill = Stat)) + geom_col(position = "dodge") + guides(fill=guide_legend(title=NULL)) + xlab("Date") + ylab("New Cases")

#covidbc.bcsubpred

```

Given this, I would predict `r round(covidbc.bcsubpred$BCCasesPredicted[53])` new cases to be announced tomorrow (March 19th)

## Total Cases and a look at South Korea

South Korea currently is showing a normal distribution for daily new cases

```{r, echo = FALSE, message= FALSE, warning= FALSE}
#ggplot(data=covidbc.data, aes(x = BCDate, y = SKCases)) +
  #geom_point()

x = covidbc.data$DayFirstCase[38:60]
y = covidbc.data$SKCases[38:60]

newdata <- data.frame(x,y)

fit <- nls(y ~ a * dnorm(x, mean, sd), data = newdata,
         start = list(mean = 50, sd = 1, a = 800))

newpred <- predict(fit, newdata = data.frame(x))
#newpred
newdata$pred <- newpred
newdata$BCDate <- covidbc.data$BCDate[38:60]

ggplot(data=newdata, aes(x = BCDate, y = y)) +
  geom_col(fill = "green") + geom_line(aes(y = pred)) + xlab("BC Date") + ylab("New Cases")

#plot(y ~ x, data = newdata)

#curve(predict(fit, newdata = data.frame(x = x)), from = 38, to = 60, add = TRUE)

```

Let's compare BC's trend with SK.

The population of Vancouver Metro is ~2.5M

The population of South Korea is ~51.5M

Which gives us a ratio of `r 2.5 / 51.5` Vancouver to South Korea

Using stats from https://www.worldometers.info/coronavirus/country/south-korea/, let's see if we can find a time that South Korea had a similar Daily New Case level.

```{r, echo=FALSE}
covidbc.data$SKCasesBC <- covidbc.data$SKCases*0.04785311832

covidbc.skdata <- subset(covidbc.data, select=c("BCDate","BCCases","SKCasesBC","BCDeaths"))

covidbc.sklong <- covidbc.skdata %>%
  gather("Stat", "Value", -BCDate)

ggplot(data=covidbc.sklong, aes(x = BCDate, y = Value, fill = Stat)) + geom_col(position = "dodge") + guides(fill=guide_legend(title=NULL)) + xlab("Date") + ylab("New Cases")
```


```{r, echo = FALSE, message= FALSE, warning= FALSE}

x = covidbc.data$DayFirstCase[38:60]
y = covidbc.data$SKCasesBC[38:60]

newdata <- data.frame(x,y)

fit <- nls(y ~ a * dnorm(x, mean, sd), data = newdata,
         start = list(mean = 50, sd = 1, a = 200))

newpred <- predict(fit, newdata = data.frame(x))
#newpred
newdata$pred <- newpred
newdata$BCDate <- covidbc.data$BCDate[38:60]
newdata$BCCases <- covidbc.data$BCCases[38:60]

ggplot(data=newdata, aes(x = BCDate, y = BCCases)) +
  geom_col(fill = "green") + geom_line(aes(y = pred)) + xlab("BC Date") + ylab("New Cases")

#plot(y ~ x, data = newdata)

#curve(predict(fit, newdata = data.frame(x = x)), from = 38, to = 60, add = TRUE)

```